---
title: "ISOM5610 Homework 2 - Blue Stem"
author: "Team 1"
date: "30 November 2018"
output: pdf_document
---

1. Data Manipulation

1.1 Check Missing Values

```{r}
setwd("~/MSBA/ISOM5610/HW2")
library(readxl)
bluestem <- as.data.frame(read_excel("bluestem.xls", sheet=1))
sum(is.na(bluestem)) #check number of missing data

bluestem[,2:9] <- lapply(bluestem[,2:9], factor) #convert to factors
colnames(bluestem)[1] <- "Sales"
colnames(bluestem)[10] <- "Index"
summary(bluestem)
str(bluestem)
```

1.2 Deseasonalize Sales

```{r}
## standardize index
bluestem$Weekday <- factor(names(bluestem[3:9])[max.col(bluestem[3:9])], levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
index_table <- unique(bluestem[names(bluestem) %in% c("Weekday", "Index")])
bluestem$Index <- bluestem$Index*7/sum(index_table[,1])
index_table <- unique(bluestem[names(bluestem) %in% c("Weekday", "Index")])

# add Deseasonalize Sales
bluestem$desSales <- bluestem$Sales/bluestem$Index
```

1.3 Add Lagged Predictors for Promotion

```{r}
library(Hmisc)
bluestem$prolag1 <- Lag(bluestem$Promotion,1)
bluestem$prolag2 <- Lag(bluestem$Promotion,2)
```

2. Data Exploration - Explore Underlying Patterns

2.1 Sales Versus Time Index

```{r}
# add the time index predictor
bluestem$timeindex <- c(1:249)

# see if there is linear trend along the time line
fit0.1 <- lm(Sales ~ timeindex,data=bluestem) # fit linear model
plot(bluestem$timeindex,bluestem$Sales, type = "l") 
# the plot is same as the time series plot:
# plot(as.ts(bluestem$Sales))
abline(fit0.1,col=2)
summary(fit0.1)
```

2.2 Deseasonalized Sales Versus Time Index

```{r}
fit0.2 <- lm(desSales ~ timeindex,data=bluestem)
plot(bluestem$timeindex,bluestem$desSales, type = "l")
# the plot is same as the time series plot:
# plot(as.ts(bluestem$desSales))
abline(fit0.2,col=2)
summary(fit0.2)
```

2.3 Nightly Indices Versus Different Days in a Week

```{r}
library(ggplot2)
library(RColorBrewer)

ggplot(index_table, aes(x = Weekday, y = Index, fill=Weekday)) + 
    geom_bar(stat = "identity") + 
    geom_text(aes(label = sprintf("%0.2f", round(Index, digits = 2))), vjust=-0.6) + 
    scale_fill_manual(name="Days of Week", values=brewer.pal(7, "Blues")[1:7]) +
    labs(y="Nightly Index", x="Days of Week")
```

2.4 Sales Versus Different Days in a Week / Promotion

```{r}
# sales vs. different days in a week
ggplot(bluestem, aes(x = Weekday, y = Sales, fill=Weekday)) + 
    geom_boxplot() + 
    geom_point(data=subset(bluestem, Promotion==1), mapping=aes(color=Promotion), size=3) + 
    scale_fill_manual(name="Days of Week", values=brewer.pal(7, "Blues")[1:7]) + 
    scale_color_manual(labels = "With Promotion", values="red") + 
    labs(x="Days of Week")

# sales with or without promotions
ggplot(bluestem, aes(x = Promotion, y = Sales, fill= Promotion)) +
    geom_boxplot()+
    scale_fill_manual(values=c(brewer.pal(7, "Blues")[4],brewer.pal(7, "Reds")[4]))
```

2.5 Deseasonalized Sales Versus Promotion

```{r}
ggplot(bluestem, aes(x = Promotion, y = desSales, fill= Promotion)) + 
    geom_boxplot() + 
    scale_fill_manual(values=c(brewer.pal(7, "Blues")[4],brewer.pal(7, "Reds")[4])) +
    labs(y="Sales/Nightly Index")
```

3. Model Building

```{r}
library(car)
# This function is a combination of 
# (1) Diagnostic plots; 
# (2) Breusch-Pagan test ("BP test" for short);
# (3) Multicollinearity check for independent variables (by VIF); and 
# (4) Cochrane-Orcutt test ("CO test" for short)
check_assumption <- function(testfit){
    #Diagnostic plots
    stdres=rstandard(testfit)
    par(mfrow=c(2,2))
    qqnorm(stdres,main='Normal Probability Plot',xlab='Normal Quantiles',ylab='Standardized Residual Quantiles')
    abline(0,1)
    plot(testfit$fitted.values,stdres,main='Versus Fits',xlab='Fitted Value',ylab='Standardized Residual')
    abline(0,0,lty=3)
    hist(stdres,main='Histogram',xlab='Standardized Residual')
    plot(as.ts(stdres),type="o",main='Versus Order',xlab='TimeIndex',ylab='Standardized Residual')
    abline(0,0,lty=3)
    #BP test
    cat("\n==========================Breusch-Pagan test==========================\n\n")
    print(ncvTest(testfit))
    #Checking multicollinearity for independent variables
    cat("\n=========Checking multicollinearity for independent variables=========\n\nVIF:\n")
    tryCatch({
         print(vif(testfit))
         },
         warning = function(msg) {
              message("Original warning message:")
              message(paste0(msg,"\n"))
              return(NULL)
         },
         error = function(msg) {
              message("Original error message:")
              message(paste0(msg,"\n"))
              return(NA)
         }
) 
    
    #Cochrane-Orcutt test
    cat("\n=========================Cochrane-Orcutt test=========================\n")
    summary(lm(testfit$residuals~Lag(testfit$residuals, 1)))
}
```

3.1 Additive Model

3.1.1 Model 0.1

```{r}
# additive model 0.1
bluestem_1 <- bluestem[,c(1:8)] 
names(bluestem_1) # a subset of bluestem with "Sales", "Promotion" dummy and seasonal dummies from ¡°Monday¡± to ¡°Saturday¡±
add.fit.1 <- lm(Sales ~ ., data = bluestem_1)
summary(add.fit.1)
add.fit.1$coefficients[2] # check promotion effect

# Checking model assumptions
check_assumption(add.fit.1)
```

3.1.2 Model 0.2 & Model 0.3 with Lagged Predictors

```{r}
# additive model 0.2 with lag 1
bluestem_1 <- bluestem[,c(1:8,13)]
names(bluestem_1) # a subset of bluestem with "Sales", "Promotion" dummy, seasonal dummies from ¡°Monday¡± to ¡°Saturday¡±, and lag 1 of promotion
add.fit.2 <- lm(Sales ~ ., data = bluestem_1)
summary(add.fit.2)
coef(summary(add.fit.2))[c(2,9),] # check promotion effect

# additive model 0.2 with lag 1 and lag 2
bluestem_1 <- bluestem[,c(1:8,13,14)]
names(bluestem_1) # a subset of bluestem with "Sales", "Promotion" dummy, seasonal dummies from ¡°Monday¡± to ¡°Saturday¡±, lag 1 and lag 2 of promotion
add.fit.3 <- lm(Sales ~ ., data = bluestem_1)
summary(add.fit.3)
coef(summary(add.fit.3))[c(2,9,10),] # check promotion effect

# check model assumptions
check_assumption(add.fit.2)
check_assumption(add.fit.3)
```

3.2 Multiplicative Model

3.2.1 Model 1.1

```{r}
# multiplicative model 1.1 with promotion dummy only
mtp.fit.1 <- lm(desSales ~ Promotion, data = bluestem)
summary(mtp.fit.1)
mtp.fit.1$coefficients[2] # check promotion effect on deseasonalized sales

# The coefficient of Promotion stands for the boost in revenues without "days of week"" effect. Such boost would be amplified by the Nightly Index for Friday / Saturday.
mtp.fit.1$coefficients[2] * index_table$Index[which(index_table$Weekday=="Friday")]
mtp.fit.1$coefficients[2] * index_table$Index[which(index_table$Weekday=="Saturday")]

# Checking model assumptions
check_assumption(mtp.fit.1)
```

3.2.2 Model 1.2 and Model 1.3 with Lagged Predictors

```{r}
# multiplicative model 1.2 with promotion dummy and lag 1
mtp.fit.2 <- lm(desSales ~ Promotion + prolag1, data = bluestem)
summary(mtp.fit.2)

# multiplicative model 1.3 with promotion dummy, lag 1 and lag 2
mtp.fit.3 <- lm(desSales ~ Promotion + prolag1 + prolag2, data = bluestem)
summary(mtp.fit.3)
```

3.2.3 Model 1.4 & Model 1.5: Using the Logged Response to Solve Autocorrelation

```{r}
# multiplicative model 1.4: try log of deseasonalized sales
bluestem$log_desSales <- log(bluestem$desSales)
mtp.fit.4 <- lm(log_desSales ~ Promotion, data = bluestem)
summary(mtp.fit.4)
coef(summary(mtp.fit.4))[2] # check promotion effect on deseasonalized sales as %
check_assumption(mtp.fit.4)


# additive model 1.5: try log of sales
bluestem$log_Sales <- log(bluestem$Sales)
bluestem_1 <- bluestem[,c(17,2:8,13)]
names(bluestem_1) # a subset of bluestem with "log_Sales", "Promotion" dummy, seasonal dummies from ¡°Monday¡± to ¡°Saturday¡±, and lag 1 of promotion
add.fit.5 <- lm(log_Sales ~ ., data = bluestem_1)
summary(add.fit.5)
coef(summary(add.fit.5))[c(2,9),] #check promotion effect on sales as %
check_assumption(add.fit.5)
```

3.3 Model Accuracy

```{r}
## 2 models are selected in first round selection: add.fit.2 and mtp.fit.1
## L1 L2 percentage---- these three criteria are used to evaluate them

Error <- matrix(0,nrow=2,ncol=3)
colnames(Error) <- c("Mean Squared Error","Absolute Error","Absolute Error Percentage")
rownames(Error) <- c("add.fit.2","mtp.fit.1")

## add.fit.2
bluestem$prolag1[1]<-0
add.pd.2 <- predict(add.fit.2,bluestem)
dif <- bluestem$Sales-add.pd.2
Error[1,1] <- crossprod(dif,dif)/249
Error[1,2] <- sum(abs(dif))/249
Error[1,3] <- sum(abs(dif/bluestem$Sales))/249

## mtp.fit.1
mtp.pd.1 <- predict(mtp.fit.1,bluestem)*bluestem$Index
dif <- bluestem$Sales-mtp.pd.1
Error[2,1] <- crossprod(dif,dif)/249
Error[2,2] <- sum(abs(dif))/249
Error[2,3] <- sum(abs(dif/bluestem$Sales))/249

knitr::kable(Error,format = "markdown")
```

3.4 Conclusion

```{r}
# selected additive model
summary(add.fit.2)
coef(summary(add.fit.2))[c(2,9),] # check promotion effect

# selected multiplicative model
summary(mtp.fit.1)
mtp.fit.1$coefficients[2] # check promotion effect on deseasonalized sales
index_table$Index[which(index_table$Weekday=="Saturday")] # standardized nightly index on Saturdays
mtp.fit.1$coefficients[2] * index_table$Index[which(index_table$Weekday=="Saturday")]
```

#######################################################
Note: The following part is not included in the report. It records our exploration of different methods to solve the autocorrelation problem including adding lagged predictors and logging the response variable, but these methods did not turn to be effective to eliminate the problem.

```{r}
## additive model with timeindex
bluestem_1_t <- bluestem[,c(1:8,15)]
names(bluestem_1_t)
add.fit.1_t <- lm(Sales ~ ., data = bluestem_1_t)
summary(add.fit.1_t)
## add lag to additive model with timeindex
# add lag 1
bluestem_1_t <- bluestem[,c(1:8,15,13)] 
names(bluestem_1_t)
add.fit.2_t <- lm(Sales ~ ., data = bluestem_1_t)
summary(add.fit.2_t)
# add lag 1 and lag 2
bluestem_1_t <- bluestem[,c(1:8,15,13,14)]
names(bluestem_1_t)
add.fit.3_t <- lm(Sales ~ ., data = bluestem_1_t)
summary(add.fit.3_t)

## multiplicative model with timeindex
mtp.fit.1_t <- lm(desSales ~ Promotion + timeindex, data = bluestem)
summary(mtp.fit.1_t)
## add lag to multiplicative model with timeindex
# add lag 1
mtp.fit.2_t <- lm(desSales ~ Promotion + timeindex + prolag1, data = bluestem)
summary(mtp.fit.2)
# add lag 1 and lag 2
mtp.fit.3_t <- lm(desSales ~ Promotion + timeindex + prolag1 + prolag2, data = bluestem)
summary(mtp.fit.3)
## try log(desSales)
mtp.fit.4_t <- lm(log_desSales ~ Promotion + timeindex, data = bluestem)
summary(mtp.fit.4_t)
mtp.fit.5_t <- lm(log_desSales ~ Promotion + timeindex + prolag1, data = bluestem)
summary(mtp.fit.5_t)
mtp.fit.6_t <- lm(log_desSales ~ Promotion + timeindex + prolag1 + prolag2, data = bluestem)
summary(mtp.fit.6_t)
```