---
title: "ISOM5610 HW2"
author: "Team 1"
date: "22 November 2018"
output: html_document
---

```{r}
setwd("~/MSBA/ISOM5610/HW2")
library(readxl)
bluestem_0 <- as.data.frame(read_excel("bluestem.xls", sheet=1))
bluestem_0[,2:9] <- lapply(bluestem_0[,2:9], factor)
colnames(bluestem_0)[1] <- "Sales"
colnames(bluestem_0)[10] <- "Index"
summary(bluestem_0)
```

#Data Exploration

```{r}
#data processing
bluestem_exp <- bluestem_0 #data for exploration
bluestem_exp$Weekday <- factor(names(bluestem_exp[3:9])[max.col(bluestem_exp[3:9])], levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
bluestem_exp$desSales <- bluestem_exp$Sales/bluestem_exp$Index

#Different Nightly Index for different Weekdays
library(ggplot2)
index_table <- unique(bluestem_exp[10:11])
ggplot(index_table, aes(x = Weekday, y = Index)) + geom_bar(stat = "identity") + geom_text(aes(label = Index), vjust=-0.6) + labs(y="Nightly Index")

#Sales on different Weekdays
ggplot(bluestem_exp, aes(x = Weekday, y = Sales)) + geom_boxplot()

#Sales with or without promotions
ggplot(bluestem_exp, aes(x = Promotion, y = Sales)) + geom_boxplot()

#De-seasonalized Sales with or without promotions
ggplot(bluestem_exp, aes(x = Promotion, y = desSales)) + geom_boxplot() + labs(y="Sales/Nightly Index")
```

## Additive Model
A regression model using dummy variables to account for the fixed Weekday??s effect.

```{r}
bluestem_1 <- bluestem_exp[1:8]
str(bluestem_1) #Sunday as the base dummy
add.fit=lm(formula = 'Sales ~ .', data = bluestem_1)
summary(add.fit)
add.fit$coefficients[2]
```
**Answer = 593.5**

```{r}
# Checking model assumptions
par(mfrow=c(2,2))
plot(add.fit)
# Checking multicollinearity for independent variables
library(HH)
vif(add.fit)
```

## Multiplicative Model
The nightly index expresses each Weekday??s effect on the popularity. 

#### 1. Use De-seasonalized Sales.

```{r}
mtp.fit=lm(formula = 'desSales ~ Promotion', data = bluestem_exp)
summary(mtp.fit)
# The coefficient of Promotion stands for the boost in revenues without Weekday effect. Such boost would be amplified by the Nightly Index for Saturday.
mtp.fit$coefficients[2] * index_table$Index[which(index_table$Weekday=="Saturday")]
```
**Answer = 801.4345**

```{r}
# Checking model assumptions
par(mfrow=c(2,2))
plot(mtp.fit)
```

#### 2. Try log(De-seasonalized Sales)

```{r}
bluestem_exp$log_desSales <- log(bluestem_exp$desSales)
mtp.fit.2=lm(formula = 'log_desSales ~ Promotion', data = bluestem_exp)
summary(mtp.fit.2)
# Checking model assumptions
par(mfrow=c(2,2))
plot(mtp.fit.2)
mtp.fit.2$coefficients[2]
# Promotion increases sales without Weekday effect by 52.284%. Now factor in the Saturday effect in terms of % increase:
mtp.fit.2$coefficients[2] * ((index_table$Index[which(index_table$Weekday=="Saturday")] - 1)/1)
```
**Answer = 97.77094%**

## Comparison

**Model assumptions:**  
The additive model is much better in terms of R-square and residual plots. The multiplicative model using log(De-seasonalized Sales) is better than the one directly using De-seasonalized Sales.  
**Business perspective:**   
The multiplicative model is better than the additive model. The additive model assumes that promotions have same boost effect on all weekdays, which is not the case. Promotional events should have better results with larger customer traffic. This is better addressed by the multiplicative model.

**huhuhuhubert**
just wanna try to modify
